Improved Approximation Algorithms for the Maximum Carpool Matching Problme

### 2017-02-15 src/intro.tex
intro

Bar-Noy et al.~\cite{bar2015improved} considered the
\textsc{Minimum $2$-Path Partition} problem.
In this problem the input is a complete graph on $3k$ vertices with
non-negative edge weights, and the goal is to partition the graph into
disjoint paths of length 2.  This problem is the special case of the
undirected carpool matching where $c(v) = 2$, for every $v \in V$.
They presented two approximation algorithms, one for the weighted case
whose is ratio is $0.5833$, and another for the unweighted case whose
ratio is $\frac{3}{4}$.

Another related problem is the \textsc{$k$-Set Packing}, where one is
given a collection of weighted sets, each containing as most $k$
elements, and the goal is to find a maximum weight subcollection of
disjoint sets.  Chandra and Halld\'orsson~\cite{chandra2001greedy}
presented a $\frac{3}{2(k+1)}$-approximation algorithm for this
problem.
%
\carpool can be seen as a special case of \textsc{$k$-Set Packing}
with $k = \cmax + 1$.  Consider a subset of nodes $U$ of size at most
$k$.  Observe that each subset of nodes has an optimal internal
assignment of passenger to drivers.  Let the weight of this assignment
be the profit of $U$, denote $p(U)$.  If $k = O(1)$, $p(U)$ can be
computed for every $U$ of size at most $k$ in polynomial time.  The
outcome is a \textsc{$k$-Set Packing} instance.  This leads to a
$\frac{3}{2(\cmax+2)}$-approximation algorithm for the case where
$\cmax = O(1)$.

@@

Section~\ref{sec:cmax} we consider \carpool with bounded maximum
capacity.   Define $\cmax \eqdf \max_{v \in V} c(v)$.
%
In Section~\ref{sec:hardness} we show that \carpool is APX-hard even
for undirected and unweighted instances with $\cmax \leq b$, for any
$b \geq 3$.
%
In Section~\ref{sec:local} we consider the unweighted variant of the
problem.  We provide another local search algorithm whose
approximation ratio is $\half + \inv{2\cmax} - \eps$, for any
$\eps>0$.  Given a parameter $k$, our algorithm starts with the empty
carpool matching.  Then, it repeatedly tries to find a better matching
by replacing $t \leq k$ arc in the current solution by $t+1$ arcs that
are not in the solution.
%
We show that our analysis is tight.  We also note that our algorithm
is a special case of the local search family that was defined
in~\cite{ACKK09} with a parameter of $2k$.  On undirected and
uncapaciated instances we have that $\cmax = \Delta$, where $\Delta$
is the maximum degree, hence the approximation ratio of our algorithm
is $\half + \inv{2\Delta} - \eps$ which is tight with respect
to~\cite{ACKK09}.

Finally, Section~\ref{sec:group} discusses the \gcp.  We first show
that the unconstrained submodular maximization formulation for
the \carpool does not work for \gcp.  We show, however, that this
problem still admits a $(\frac{1}{2} -\varepsilon)$-approximation
algorithm by extending our first the local search algorithm.  In
addition, we show that the second local search algorithm generalizes
to unweighted \gcp with the same approximation ratio.

>>>>

Bar-Noy et al.~\cite{bar2015improved} considered the
\textsc{Minimum $2$-Path Partition} problem.
In this problem the input is a complete graph on $3k$ vertices with
non-negative edge weights, and the goal is to partition the graph into
disjoint paths of length 2.  This problem is the special case of the
undirected carpool matching where $c(v) = 2$, for every $v \in V$.
They presented two approximation algorithms, one for the weighted case
whose ratio is $0.5833$, and another for the unweighted case whose
ratio is $\frac{3}{4}$.

Another related problem is the \textsc{$k$-Set Packing}, where one is
given a collection of weighted sets, each containing at most $k$
elements, and the goal is to find a maximum weight subcollection of
disjoint sets.  Chandra and Halld\'orsson~\cite{chandra2001greedy}
presented a $\frac{3}{2(k+1)}$-approximation algorithm for this
problem.
%
\carpool can be seen as a special case of \textsc{$k$-Set Packing}
with $k = \cmax + 1$.  Consider a subset of nodes $U$ of size at most
$k$.  Observe that each subset of nodes has an optimal internal
assignment of passenger to drivers.  Let the weight of this assignment
be the profit of $U$, denote $p(U)$.  If $k = O(1)$, $p(U)$ can be
computed for every $U$ of size at most $k$ in polynomial time.  The
outcome is a \textsc{$k$-Set Packing} instance.  This leads to a
$\frac{3}{2(\cmax+2)}$-approximation algorithm for the case where
$\cmax = O(1)$.

@@

In Section~\ref{sec:cmax} we consider \carpool with bounded maximum
capacity.   Define $\cmax \eqdf \max_{v \in V} c(v)$.
%
In Section~\ref{sec:hardness} we show that \carpool is APX-hard even
for undirected and unweighted instances with $\cmax \leq b$, for any
$b \geq 3$.
%
In Section~\ref{sec:local} we consider the unweighted variant of the
problem.  We provide another local search algorithm whose
approximation ratio is $\half + \inv{2\cmax} - \eps$, for any
$\eps>0$.  Given a parameter $k$, our algorithm starts with the empty
carpool matching.  Then, it repeatedly tries to find a better matching
by replacing $t \leq k$ arc in the current solution by $t+1$ arcs that
are not in the solution.
%
We show that our analysis is tight.  We also note that our algorithm
is a special case of the local search family that was defined
in~\cite{ACKK09} with a parameter of $2k$.  On undirected and
uncapaciated instances we have that $\cmax = \Delta$, where $\Delta$
is the maximum degree, hence the approximation ratio of our algorithm
is $\half + \inv{2\Delta} - \eps$ which is tight with respect
to~\cite{ACKK09}.

Finally, Section~\ref{sec:group} discusses the \gcp.  We first show
that the unconstrained submodular maximization formulation for
the \carpool does not work for \gcp.  We show, however, that this
problem still admits a $(\frac{1}{2} -\varepsilon)$-approximation
algorithm by extending our first local search algorithm.  In
addition, we show that the second local search algorithm generalizes
to unweighted \gcp with the same approximation ratio.
